To be FAIR: Theory Specification Needs an Update

### 2025-02-19 manuscript.Rmd
update 2025-02-19 10:53:20.811478

At present, there are several impediments to theories' accessibility.
To the extent that theories are still contained within papers,
paywalls erected by commercial publishers constitute a barrier.
Open Access publishing thus increases the accessibility of all academic output, including theory.
A second impediment is more indirect:
While open access publishing increases practical access to theories,
accessibility also requires clear and explicit communication.
<!-- NvD: I am not sure if we need to go into this much detail as presented in the rest of this Section. My suggestion is to remove most of it. I'll indicate which elements I suggest keeping -->
This property of good theories has been dubbed "discursive survival [...], the ability to be understood" [@guestWhatMakesGood2024]. 
At present, psychological theories are often ambiguous, rendering them difficult to understand [@frankenhuisStrategicAmbiguitySocial2023].
It is important to acknowledge the *indeterminacy of translation* [@quineReasonsIndeterminacyTranslation1970]:
which holds that every communicative utterance has multiple alternative translations, with no *objective* means of choosing the correct one.
It follows that an idea cannot be formalized to the point that it becomes unambiguously interpretable.
This places a theoretical upper bound on theories' ability to be understood.

>>>>

At present, there are several impediments to theories' accessibility.
To the extent that theories are still contained within papers,
paywalls erected by commercial publishers constitute a barrier.
Open Access publishing thus increases the accessibility of all academic output, including theory.
A second impediment is more indirect:
While open access publishing increases practical access to theories,
accessibility also requires clear and explicit communication.
<!-- NvD: I am not sure if we need to go into this much detail as presented in the rest of this Section. My suggestion is to remove most of it. I'll indicate which elements I suggest keeping -->
This property of good theories has been dubbed "discursive survival [...], the ability to be understood" [@guestWhatMakesGood2024]. 
At present, psychological theories are often ambiguous, rendering them difficult to understand [@frankenhuisStrategicAmbiguitySocial2023].
It is important to acknowledge the *indeterminacy of translation* [@quineReasonsIndeterminacyTranslation1970]:
which holds that every communicative utterance (e.g., a statement in natural language, a mathematical formula, et cetera)
has multiple alternative translations, with no *objective* means of choosing the correct one.
It follows that an idea cannot be formalized to the point that it becomes unambiguously interpretable.
This places a theoretical upper bound on theories' ability to be understood.
